Controlled Experiments for Word Embeddings

### 2015-12-14 article/main.tex
incorporate adriaan corrections to second version

Consider first the word vectors for the pseudowords associated to the word
\word{the}.  Notice that the cosine similarity of the word vectors for
\word{THE\_1} and \word{THE\_i} decreases monotonically with $i$, while the
cosine similarity of the word vectors for \word{THE\_i} and \word{THE\_18}
increases monotonically with $i$.  Indeed the direction of
the word vector \word{THE\_i} changes systematically, interpolating between the
directions of the word vectors of the highest-frequency pseudoword
\word{THE\_1} and the lowest-frequency pseudoword \word{THE\_18}.  The same
trend is apparent (though over shorter frequency ranges) for all the families of pseudowords
other than that for \word{VOID}.

Consider now the word vectors for pseudowords derived from the meaningless word
\word{VOID}.  The word vectors for \word{VOID\_7}, \dots, \word{VOID\_13}
are approximately orthogonal to one another, just as would be expected from
randomly drawn vectors in a high dimensional space.
As the pseudoword \word{VOID} occurs by
construction in every context, a much higher number of samples is
required to capture its co-occurrence distribution, and
thereby to learn its word vector (the same is true, but to a lesser
extent, for the stopword \word{the}).
We conclude that the vectors corresponding to the lower frequency
pseudowords \word{VOID\_7}, \dots, \word{VOID\_13} have not been trained on a
sufficient number of samples to establish their proper direction.
These vectors are excluded from further analysis.
The vectors for \word{VOID\_1}, \dots, \word{VOID\_6}, on the other hand,
exhibit the smooth change in word vector direction with word frequency described in
the previous paragraph.

>>>>

Consider first the vectors for the pseudowords associated to the word
\word{the}.  Notice that the cosine similarity of the vectors for
\word{THE\_1} and \word{THE\_i} decreases monotonically with $i$, while the
cosine similarity of the vectors for \word{THE\_i} and \word{THE\_18}
increases monotonically with $i$.  Indeed the direction of
the vector \word{THE\_i} changes systematically, interpolating between the
directions of the vectors of the highest-frequency pseudoword
\word{THE\_1} and the lowest-frequency pseudoword \word{THE\_18}.  The same
trend is apparent (though over shorter frequency ranges) for all the families of pseudowords
other than that for \word{VOID}.

Consider now the vectors for pseudowords derived from the meaningless word
\word{VOID}.  The vectors for \word{VOID\_7}, \dots, \word{VOID\_13}
are approximately orthogonal to one another, just as would be expected from
randomly drawn vectors in a high dimensional space.
As the pseudoword \word{VOID} occurs by
construction in every context, a much higher number of samples is
required to capture its co-occurrence distribution, and
thereby to learn its vector (the same is true, but to a lesser
extent, for the stopword \word{the}).
We conclude that the vectors corresponding to the lower frequency
pseudowords \word{VOID\_7}, \dots, \word{VOID\_13} have not been trained on a
sufficient number of samples to establish their proper direction.
These vectors are excluded from further analysis.
The vectors for \word{VOID\_1}, \dots, \word{VOID\_6}, on the other hand,
exhibit the smooth change in vector direction with word frequency described in
the previous paragraph.
